### 2015-01-06 sharp-sigma-one-plus.tex
6 ene

Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on. As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts. Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
\begin{multline*}
\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
\end{multline*}
where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
\begin{multline*}
\psi_i(\X) = \exists\u_1\cdots\exists\u_n\bigg[\alpha^j_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
\end{multline*}
Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
\begin{multline}
\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\varphi_1(\v,\X)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{i-1}(\v,\X)\vee\psi_i(\X)).
\end{multline}
Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$. Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}. As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
\begin{enumerate}
\item[a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:v.$
\item[b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assigment to $\X$ for any of these formulas. (2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$. However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
\item[c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (1) and (2).
\end{enumerate}
We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$.

>>>>

Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on. As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts. Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
\begin{multline*}
\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
\end{multline*}
where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
\begin{multline*}
\psi_i(\X) = \forall\x\:\neg\alpha(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
\end{multline*}
Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
\begin{multline}
\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\varphi_1(\v,\X)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{i-1}(\v,\X)\vee\psi_i(\X)).
\end{multline}
Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$. Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula. As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
\begin{enumerate}
\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$. (2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$. However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\exists\x\:\psi_q(\P)$, so necessarily $\A\models\exists\x\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q.$
\end{enumerate}
